### 2020-04-28 reports/minority-supplement-figures.tex
made minority diff heatmap for K=10, M=4; added to results.tex

\begin{figure}[H]
  \centering
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p00.pdf}
  \end{subfigure}
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p05.pdf}
  \end{subfigure}
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p10.pdf}
  \end{subfigure} \\
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p15.pdf}
  \end{subfigure}
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p20.pdf}
  \end{subfigure}
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p25.pdf}
  \end{subfigure}
  \\
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p30.pdf}
  \end{subfigure}
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p35.pdf}
  \end{subfigure}
  \begin{subfigure}{0.315\textwidth}
    \includegraphics[width=\textwidth]{Figures/full-minority-homophily=0p40.pdf}
  \end{subfigure} \\
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=0.7\textwidth]{Figures/full-minority-homophily=0p45.pdf}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \includegraphics[width=0.7\textwidth]{Figures/full-minority-homophily=0p50.pdf}
  \end{subfigure}
  \caption{}
  \label{fig:}
\end{figure}